---
title: "HW 5"
author: "Clement Mugenzi"
date: "11/6/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(skimr)
library(purrr)
library(tidyr)
library(rvest)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "right"))
```

# Problem 1

This is the code chunk that introduces missing data in the **iris** dataset.

```{r}
set.seed(10)

iris_with_missing = iris %>% 
  map_df(~replace(.x, sample(1:150, 20), NA)) %>%
  mutate(Species = as.character(Species)) %>% view()
```

## Writing the Function

I will then create a function that will replace missing numeric values by the mean of that particular variable and also this function will replace missing character values with the the species **virginica**.


```{r}
missingness = function(x) {
  if (!is.numeric(x)) {
    replace(x, is.na(x), "virginica")
  } else if (is.numeric(x)) {
  replace(x, is.na(x), mean(x, na.rm = TRUE))
  }
}
```

```{r}
iris1 = 
  map_df(.x = iris_with_missing, ~ missingness(.x)) %>% 
  knitr::kable()
iris1
```


# Problem 2

Reading in data and tidying it.

```{r, message=FALSE}
data_path = "data/" 
files = dir(data_path, pattern = "*.csv") 

data = data_frame(filename = files) %>% 
  mutate(file_contents = map(filename, ~ read_csv(file.path(data_path, .))))  
data = data %>% 
  unnest() %>% 
  separate(col = filename, into = c("arm", "id")) %>% 
  mutate(
    arm = recode(arm, "con" = "Control", "exp" = "Experimental")) %>% 
  select(id, arm, week_1:week_8) %>% view()
```

Next, I will build the spaghetting plot showing observations of each subject overtime.

```{r}
data %>%
  pivot_longer(
    week_1:week_8,
    names_to = "week",
    values_to = "observation") %>% 
  ggplot(aes(x = week, y = observation, color = id, group = id)) +
  geom_line() +
  facet_grid(~arm) +
  labs(
    title = "Observation of each Subject Overtime",
    x = "Week",
    y = "Observations")
  
```

As we move from week_1 to week_28, there is an overall plateau trend (despite within groups fluctuations) in observations of the control group (suggesting a placebo was administered) and an overall increase in observations for the intervantion group. Since we are dealing with a longitidinal study, I suspect the treatment for the variable being studied here is really efficient (or inefficient) and can safely conclude that there is a significant difference between the intervention and control groups at baseline (week_1) compared to the same groups after the treatment (week_8).








































































